Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-09 csf-2017/paper.tex
changed f1 and f2 text in example section

\section{Examples}

\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $r_\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

>>>>

\section{Examples}
In this section, we display how capability safety allows us to prove
local state encapsulation and control-flow integrity properties. For
presentation purposes, we do not formally state the lemmas we
prove. The interested reader can find all the technical details in the
techincal report.

\subsection{Encapsulation of local state}
The capability machine provides encapsulation of local state. The program \texttt{f1} in \figurename~\ref{fig:prog-f1} and the program \texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.

The programs \texttt{f1} and \texttt{f2} are very similar. They both
store some local data, call an untrusted piece of code, and finally
make sure that their local state is unchanged. They differ in the way
they do this. \texttt{f1} assumes it has access to a stack capability
and uses the stack to store its local state. It also uses the
stack-based calling convention to call the untrusted code. On the
other hand, \texttt{f2} allocates a new piece of memory for its local
state and uses an activation-record based calling convention
(described in the technical report) to run the adveserial code.

For both programs, we can prove that if they are linked up with an
adversary that is allowed to allocate new memory but otherwise has no
capabilities, then executing the program, the assertion will never
fail.
We can only prove this, because the capability machine encapsulates
the local state of the two programs. The two examples also illustrate
the versatility of the logical relation. The two examples use
different calling convention, but the logical relation is not specific
to any calling convention, so we can use it to reason about both
programs. In order to reason about the unknown untrusted part, we use
the Fundamental Theorem to argue that it must be capability safe.
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f1: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[]$)
    pop $r_1$
    assert $r_1$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a stack-based calling convention.}
  \label{fig:prog-f1}
\end{figure}

\todo[inline]{give a taste of a proof by giving high-level proof here?
Maybe later for one of the other examples}

@@

\subsection{Well-bracketed control-flow}

The program $f3$ in \figurename~\ref{fig:prog-f3}
\begin{figure}[htbp]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}
